Precise predictions for same-sign W-bosons scattering at the LHC

### 2018-02-16 Draft/introduction.tex
updated figures with PDF uncertainties fixed, uptated shower.tex implemented suggestions by PG in the intro

Vector-boson scattering (VBS) is a class of processes that make it possible 
to probe the nature of Higgs-vector-vector couplings and triple and quartic gauge couplings.
It is usually understood that VBS refers to the scattering of massive vector bosons ($\PW^\pm,\PZ$), which therefore couple to the Higgs boson and can be longitudinally polarised.
The scattering of longitudinally-polarised bosons is of particular interest, because the corresponding matrix elements feature both gauge and unitarity cancellations that strongly depend on the actual structure of the Higgs sector.
A detailed study of this class of processes will therefore further constrain the Higgs couplings and hint at or exclude non-Standard Model Higgs bosons.

The process with two positively charged $\PW$ bosons is the VBS process with the largest signal-to-background ratio at the LHC, for which evidence has been found in $8 \TeV$ data~\cite{Aad:2014zda,Khachatryan:2014sta} and has been subsequently
observed~\cite{Sirunyan:2017ret} and measured~\cite{Aaboud:2016ffv} in $13\TeV$ data. At present 
the measurements of VBS processes are limited by statistics but the situation will change in a near future.
On the theoretical side, it is thus of prime importance to provide precise predictions and infer their related systematic errors.

The $\PW^+\PW^+$ scattering is the simplest VBS process to calculate, because of the double-charge structure of the leptonic final state that limits the number of partonic processes and total number of Feynman diagrams for each process.
It also implies that irreducible backgrounds are comparatively small, which make this VBS process experimentally favourable in comparison to \eg $\PW^+\PW^-$ scattering, which has the largest cross section.
Therefore, the $\PW^+\PW^+$ scattering is an ideal candidate for a detailed study of various theoretical predictions.

@@

The hadronic process ito be considered is $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$, which includes the $\PW^+\PW^+$ scattering.
This final state possesses three contributions at leading order (LO) whose coupling orders are $\mathcal{O}{\left(\alpha^{6}\right)}$, $\mathcal{O}{\left(\alpha_{\rm s}\alpha^{5}\right)}$, and $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$.
They are commonly referred to as electroweak (EW), interference, and QCD contributions, respectively.%
\footnote{The EW contribution is sometimes referred to as the VBS contribution even it possesses non-VBS contributions.}
Therefore, the present work starts with a LO study of these three contributions as a function of typical VBS cuts.
This allows to quantify the various contributions to the final state $\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj$.
This is followed by a LO comparison between the various predictions at the level of the cross section and differential distributions.

@@

The article is organised as follows:
First, we define the process under study in Sec.~\ref{sec:definition}.
Various approximation at LO and NLO are described in Sec.~\ref{sec:details}.
This is followed by a presentation of the programs used for the computations.
Sections~\ref{sec:LO} and \ref{sec:NLO} are devoted to a LO and NLO study at fixed order, respectively.
Section~\ref{sec:matching} then adds the effect of parton showers and hadronization to the fixed-order results by matching the two.
The last section consists in concluding remarks and recommendations for experimental collaborations.

>>>>

Vector-boson scattering (VBS) at a hadron collider 
usually refers to the interaction of massive vector-bosons ($\PW^\pm,\PZ$),
irradiated by partons (quarks) of the incoming protons, 
which in turn are deflected from the beam direction 
and enter the volume of the particle detectors.
As a consequence, the typical signature of VBS events
is characterised by two energetic jets 
and four fermions,
originating from the decay of the two vector bosons.
Among the possible diagrams,
the scattering process can be mediated also by a Higgs boson
and involve in particular their longitudinal component.
The interaction of of longitudinally polarised bosons is of particular interest, 
because the corresponding matrix elements feature both gauge and unitarity cancellations 
that strongly depend on the actual structure of the Higgs sector.
A detailed study of this class of processes will therefore further constrain the Higgs couplings 
at a very different energy scale with respect to the Higgs boson mass,
and hint at, or exclude, non-Standard Model behaviours.

The VBS process involving two same-sign $\PW$ bosons has the largest signal-to-background ratio at the LHC:
evidence for it was found at the centre-of-mass energy of $8~\TeV$ already~\cite{Aad:2014zda,Khachatryan:2014sta},
and it has been recently observed~\cite{Sirunyan:2017ret} and measured~\cite{Aaboud:2016ffv} 
at $13\TeV$ as well.
Presently, the measurements of VBS processes are limited by statistics, but the situation will change in a near future.
On the theoretical side, 
it is thus of prime importance to provide predictions with systematic uncertainties
at least comparable to the current and experimental precisions~\cite{CMS:2016rcn}.

The $\PW^+\PW^+$ scattering is also the simplest VBS process to calculate, 
because the double-charge structure of the leptonic final state 
limits the number of partonic processes and total number of Feynman diagrams for each process.
Therefore, this process is the ideal candidate for a comparative study of the different simulation tools.

@@

The full gauge-invariant process including the $\PW^+\PW^+$ scattering 
 is $\Pp\Pp\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj+\mathrm{X}$.
This final state receives three contributions at leading order (LO) whose coupling orders are $\mathcal{O}{\left(\alpha^{6}\right)}$, $\mathcal{O}{\left(\alpha_{\rm s}\alpha^{5}\right)}$, and $\mathcal{O}{\left(\alphas^{2}\alpha^{4}\right)}$.
They are commonly referred to as electroweak (EW), interference, and QCD contributions, respectively.%
\footnote{The EW contribution is sometimes referred to as the VBS contribution, even it involves also non-VBS contributions.}
Therefore, the present work starts with a LO study of these three contributions as a function of typical VBS cuts.
This allows to quantify the various contributions to the final state $\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}\,\Pj\Pj$.
This is followed by a LO comparison between the various predictions at the level of the cross section and differential distributions.

@@

The article is organised as follows:
first, we define the process under study in Sec.~\ref{sec:definition}.
The various approximation which are provided by the different computer codes at LO and NLO are described in Sec.~\ref{sec:details}.
This is followed by a presentation of the programs used for the computations.
Sections~\ref{sec:LO} and \ref{sec:NLO} are devoted to a LO and NLO study at fixed order, respectively.
Section~\ref{sec:matching} complements our study by comparing predictions at LO and LO which include the effect of parton showers and hadronization.
The last section consists in concluding remarks and recommendations for experimental collaborations.

### 2018-02-16 Draft/shower.tex
updated figures with PDF uncertainties fixed, uptated shower.tex implemented suggestions by PG in the intro

\begin{table}[h!]
    \centering
    \begin{tabular}{c|r@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline\hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.450 ^{+2\%}_{-1\%} {}^{+5\%}_{-6\%} $ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}&  $1.445 $ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Pythia8}, $\Gamma_{\rm resc}$&  $1.368$ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}, $\Gamma_{\rm resc}$&  $1.363$ & $0.004$  \\
        {\sc POWHEG}  &  $1.3642$ & $0.0004$  \\
        {\sc VBFNLO}+{\sc Herwig7-Dipole} &  $1.3389 ^{+0\%}_{-1\%}$ & $0.0006$  \\
        {\sc VBFNLO}+{\sc Herwig7-Default} &  $1.3067$ & $0.0006$  \\
    \end{tabular}
    \caption{\label{tab:PSratesNLO} Rates at NLO-QCD  accuracy matched to parton shower within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process. The {\sc MG5\_aMC} results with $\Gamma_{\rm resc}$ 
    are rescaled to account for the effect related to the boson widths computed by {\sc MadSpin} (see the text for details). For
    {\sc VBFNLO}+{\sc Herwig7-Dipole}, the three-point scale uncertainties is shown, while for  {\sc MG5\_aMC}+{\sc Pythia8} the two displayed uncertainties
are respectively the nine-point scale uncertainty and the PDF one.}
\end{table}

We now present the results of predictions matched to parton shower: the total rates within VBS cuts are displayed in Tables~\ref{tab:PSratesLO} and
\ref{tab:PSratesNLO}, at LO and NLO
accuracy respectively. For {\sc MG5\_aMC}, 
the numbers whith $\Gamma_{\rm resc}$ are rescaled to 
take into account the width effects described in the above paragraph.d  At NLO accuracy, for {\sc MG5\_aMC} + {\sc Pythia8} and {\sc VBFNLO}-{\sc Dipole}, we also quote
theoretical uncertainties: for the former, we show both PDF and scale uncertainties, obtained via exact reweighting~\cite{Frederix:2011ss} by varying independently the renormalization and factorization
scale by a factor 2 around its central value, Eq.~\ref{eq:defscale} (nine-point variations); for the latter, we show the 
three-point scale uncertainties, obtained by considering correlated variations of the renormalization and factorization scales. 
We observe that, once the width effect is taken into
account, total rates from different tools agree within few per cent, both at LO and NLO. Larger discrepancies however will appear for differential observables, which we are going to discuss in
the following. Before turning to the differential observables, we point out the smallness of the theory uncertainties due to the scale variations, both when
scales are varied independently and when they are varied in a correlated manner. On the other hand PDF uncertainties are larger, at the level of 
$\pm 5\%$. Concerning differential kinematic distributions, for each observable we will display results two plots, shown side-by-side. In the plot on the left (right), (N)LO+PS predictions are shown
with different colours in the main frame. In the inset, these predictions are compared in both cases with a fixed-order prediction at NLO accuracy (obtained with
{\sc VBFNLO} \emph{i.e.}\ the VBS approximation with $s$-channel contributions). For the differential observables, the {\sc MG5\_aMC} predictions are \emph{not} rescaled to compensate for the width effect mentioned above. As for the table, we show theoretical uncertainties for the NLO+PS samples
obtained with {\sc VBFNLO} and {\sc MG5\_aMC}: 
again, for the first the band corresponds to three-point variations, while for the second the darker (lighter) band corresponds to nine-point 
scale variations (plus PDF uncertainties, linearly added).

>>>>

\begin{table}[h!]
    \centering
    \begin{tabular}{c|r@{ $\pm$ }l}
      Code  &  \multicolumn{2}{c}{$\sigma[\rm{fb}]$}  \\
        \hline\hline
        {\sc MG5\_aMC}+{\sc Pythia8}&  $1.450 ^{+2\%}_{-1\%} {}^{+2\%}_{-2\%} $ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}&  $1.445 $ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Pythia8}, $\Gamma_{\rm resc}$&  $1.368$ & $0.004$  \\
        {\sc MG5\_aMC}+{\sc Herwig++}, $\Gamma_{\rm resc}$&  $1.363$ & $0.004$  \\
        {\sc POWHEG}  &  $1.3642$ & $0.0004$  \\
        {\sc VBFNLO}+{\sc Herwig7-Dipole} &  $1.3389 ^{+0\%}_{-1\%}$ & $0.0006$  \\
        {\sc VBFNLO}+{\sc Herwig7-Default} &  $1.3067$ & $0.0006$  \\
    \end{tabular}
    \caption{\label{tab:PSratesNLO} Rates at NLO-QCD  accuracy matched to parton shower within VBS cuts obtained with the different codes used in this comparison,
    for the ${\rm p}{\rm p}\to\mu^+\nu_\mu{\rm e}^+\nu_{\rm e}{\rm j}{\rm j}$ process. The {\sc MG5\_aMC} results with $\Gamma_{\rm resc}$ 
    are rescaled to account for the effect related to the boson widths computed by {\sc MadSpin} (see the text for details). For
    {\sc VBFNLO}+{\sc Herwig7-Dipole}, the three-point scale uncertainties is shown, while for  {\sc MG5\_aMC}+{\sc Pythia8} the two displayed uncertainties
are respectively the nine-point scale uncertainty and the PDF one.}
\end{table}

We now present the results of predictions matched to parton shower: the total rates within VBS cuts are displayed in Tables~\ref{tab:PSratesLO} and
\ref{tab:PSratesNLO}, at LO and NLO
accuracy respectively. For {\sc MG5\_aMC}, 
the numbers whith $\Gamma_{\rm resc}$ are rescaled to 
take into account the width effects described in the above paragraph.d  At NLO accuracy, for {\sc MG5\_aMC} + {\sc Pythia8} and {\sc VBFNLO}-{\sc Dipole}, we also quote
theoretical uncertainties: for the former, we show both PDF and scale uncertainties, obtained via exact reweighting~\cite{Frederix:2011ss} by varying independently the renormalization and factorization
scale by a factor 2 around its central value, Eq.~\ref{eq:defscale} (nine-point variations); for the latter, we show the 
three-point scale uncertainties, obtained by considering correlated variations of the renormalization and factorization scales. 
We observe that, once the width effect is taken into
account, total rates from different tools agree within few per cent, both at LO and NLO. Larger discrepancies however will appear for differential observables, which we are going to discuss in
the following. Before turning to the differential observables, we point out the smallness of the theory uncertainties due to the scale variations, both when
scales are varied independently and when they are varied in a correlated manner, as well as those due to PDF.  
Concerning differential kinematic distributions, for each observable we will display results two plots, shown side-by-side. In the plot on the left (right), (N)LO+PS predictions are shown
with different colours in the main frame. In the inset, these predictions are compared in both cases with a fixed-order prediction at NLO accuracy (obtained with
{\sc VBFNLO} \emph{i.e.}\ the VBS approximation with $s$-channel contributions). For the differential observables, the {\sc MG5\_aMC} predictions are \emph{not} rescaled to compensate for the width effect mentioned above. As for the table, we show theoretical uncertainties for the NLO+PS samples
obtained with {\sc VBFNLO} and {\sc MG5\_aMC}: 
again, for the first the band corresponds to three-point variations, while for the second the darker (lighter) band corresponds to nine-point 
scale variations (plus PDF uncertainties, linearly added).
